---
title: "Managing Assets and Tokens"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Managing Assets and Tokens}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Introduction

The `multichainr` package provides a high-level R interface to the MultiChain JSON-RPC API. This vignette demonstrates the lifecycle of managing both **fungible assets** (like currencies or loyalty points) and **non-fungible tokens (NFTs)**.

To follow this guide, ensure you have the MultiChain binaries (`multichaind` and `multichain-util`) installed on your system.

```{r setup}
library(multichainr)

# Set the path to your MultiChain binaries
mc_set_path(Sys.getenv("MULTICHAIN_PATH"))
```

## 1. Node Initialization

MultiChain operates as a permissioned blockchain. We start by initializing a new chain and launching the daemon.

```{r init}
chain_name <- "asset_demo_chain"

# Create the blockchain configuration
mc_node_init(chain_name)

# Start the node in the background
mc_node_start(chain_name)

# Allow the node a few seconds to initialize the wallet and network
Sys.sleep(3) 

# Connect to the local node
config <- mc_get_config(chain_name)
conn <- mc_connect(config)
```

## 2. Permissions and Addresses

By default, MultiChain is secure. Addresses must be granted specific permissions to issue or receive assets.

```{r addresses}
# Generate new wallet addresses
issuer_addr <- mc_get_new_address(conn)
recipient_addr <- mc_get_new_address(conn)

# Grant 'issue' and 'send' permissions to the issuer
mc_grant(conn, issuer_addr, "issue,send,receive")

# Grant 'receive' permissions to the recipient
mc_grant(conn, recipient_addr, "receive")
```

## 3. Fungible Assets

Fungible assets are divisible and interchangeable. We can create an "open" asset to allow the supply to be increased later.

```{r fungible}
# Define an asset named 'gold' with 2 decimal places (units = 0.01)
gold_params <- list(name = "gold", open = TRUE)

# Initial issuance of 1,000 units
mc_issue(conn, issuer_addr, gold_params, quantity = 1000, units = 0.01)

# Increase the total supply by an additional 500 units
mc_issue_more(conn, issuer_addr, "gold", 500)

# Verify the balance of the issuer
issuer_balances <- mc_get_address_balances(conn, issuer_addr)
print(issuer_balances)
```

## 4. Non-Fungible Tokens (NFTs)

In MultiChain 2.x, NFTs are managed as unique tokens within a non-fungible "Parent Asset."

### Creating the NFT Parent
To support individual tokens, the parent asset must be created with `fungible = FALSE` and an initial quantity of `0`.

```{r nft_parent}
art_params <- list(
  name = "art", 
  fungible = FALSE, 
  open = TRUE
)

# Issue the parent container
mc_issue(conn, address = issuer_addr, name = art_params, quantity = 0, units = 1)
```

### Issuing Unique Tokens
Once the parent exists, we can issue individual tokens (NFTs). Each token can include custom JSON metadata.

```{r issue_tokens}
# Issue a unique token "painting_001" with specific metadata
mc_issue_token(conn,
               address = issuer_addr, 
               asset = "art", 
               token = "painting_001", 
               quantity = 1, 
               token_details = list(artist = "Leonardo da Vinci", year = 1503))

# Retrieve specific token metadata
token_info <- mc_get_token_info(conn, "art", "painting_001")
print(token_info)
```

## 5. Transfers and Multibalances

We can transfer assets using specific convenience functions or the general-purpose `mc_send` function for complex transactions (like NFTs).

```{r transfers}
# 1. Send 100 units of fungible 'gold'
mc_send_asset(conn, recipient_addr, "gold", 100)

# 2. Send the NFT 'painting_001'
# NFTs require a nested list structure specifying the parent and the token name
nft_transfer <- list(
  art = list(
    token = "painting_001",
    qty = 1
  )
)
mc_send(conn, recipient_addr, nft_transfer)

# 3. Inspect the recipient's token balances
# The output is a clean data frame with an 'address' column
recipient_tokens <- mc_get_token_balances(conn, recipient_addr)
print(recipient_tokens)
```

## 6. Node Shutdown and Cleanup

When the work is complete, it is important to stop the node and, if necessary, remove the data directory.

```{r cleanup}
mc_node_stop(conn)
Sys.sleep(2)

# Determine data directory for cleanup
if (.Platform$OS.type == "windows") {
  base_dir <- file.path(Sys.getenv("APPDATA"), "MultiChain")
} else if (Sys.info()["sysname"] == "Darwin") {
  base_dir <- file.path(Sys.getenv("HOME"), "Library/Application Support/MultiChain")
} else {
  base_dir <- file.path(Sys.getenv("HOME"), ".multichain")
}

chain_dir <- file.path(base_dir, chain_name)
if (dir.exists(chain_dir)) unlink(chain_dir, recursive = TRUE)
```

## Summary

In this vignette, we demonstrated how to:

1. Configure permissions using `mc_grant`.
2. Issue open fungible assets using `mc_issue` and `mc_issue_more`.
3. Set up a non-fungible environment by setting `fungible = FALSE`.
4. Issue and query unique tokens (NFTs) via `mc_issue_token` and `mc_get_token_info`.
5. Perform multi-asset transfers and monitor balances.
